---
title: "Multi-Format Quarto Demo"
author: "Gordon Wright"
date: today
format:
  revealjs: 
    output-ext: .htm
  html:
    toc: true
    theme: cosmo
    self-contained: true
  pdf:
    toc: true
    number-sections: true
  docx:
    toc: true
    number-sections: true
execute:
  echo: true
  warning: false
---

## Introduction

This document demonstrates a single Quarto source rendering to multiple formats.
The same `.qmd` file produces HTML, PDF, RevealJS (presentations) and Word output without modification.

## A Simple Calculation

Here is an inline calculation: the mean of 1 to 10 is `r mean(1:10)`.

```{r}
#| label: summary-stats
#| fig-cap: "Distribution of a normal sample"

set.seed(42)
x <- rnorm(200, mean = 50, sd = 10)
summary(x)
```

## A Plot

```{r}
#| label: fig-histogram

hist(
  x,
  main   = "Sample Distribution",
  xlab   = "Value",
  col    = "steelblue",
  border = "white"
)
```

## A Table

```{r}
#| label: tbl-iris
#| tbl-cap: "First six rows of the iris dataset"

knitr::kable(head(iris))
```

## Cross-referencing

See @fig-histogram for the distribution and @tbl-iris for a sample table.
Quarto handles cross-references consistently across all output formats.

## Maths

Inline: $\bar{x} = \frac{1}{n}\sum_{i=1}^{n} x_i$

Display:

$$
\sigma^2 = \frac{1}{n-1}\sum_{i=1}^{n}(x_i - \bar{x})^2
$$

## Callouts

::: {.callout-note}
Callouts render in HTML and PDF. In Word output they appear as styled blockquotes.
:::

::: {.callout-warning}
PDF rendering requires a LaTeX engine. Install TinyTeX with:
`quarto install tinytex`
:::

## Rendering

Render all formats at once from the terminal:

```bash
quarto render demo.qmd
```

Or a specific format:

```bash
quarto render demo.qmd --to pdf
quarto render demo.qmd --to html
quarto render demo.qmd --to docx
```
